On the Bit Complexity of Finding Points in Connected Components of a Smooth Real Hypersurface

### 2020-02-05 main.tex
Updates from Overleaf

\keywords{Real algebraic geometry, transversality, Noether
  normalization, bit complexity}

@@

Let $f \in \ZZ[X_1,\hdots,X_n]$ have total degree $d$ with $V(f) \subset \C^n$ smooth. 
%
\begin{theorem}
There exists a hypersurface $\Delta_1 \subset \C^{n^2}$ of degree at most $d^{2n^2}$ with the property that, if $A \in \C^{n^2} - \Delta_1$ then, for each $i \in \{1,\hdots,n\},$ $f^A$ satisfies $\textbf{H}_i(1)$.
\end{theorem}

For each $i \in \{1,\hdots,n\}$, let $\Phi_i$ be the polynomial mapping 
\[
\Phi_i : (\xb,A) \mapsto \left(f^A(\xb), \frac{\partial f^A}{\partial X_{i+1}}(\xb),\hdots,\frac{\partial f^A}{\partial X_n}(\xb)\right),
\] 
so that $\Phi_{i,A}^{-1}(0)= W(\pi_i,V^A).$
%
\begin{proposition} 
$\bz$ is a regular value of $\Phi_i$.
\end{proposition}
%
\begin{proof}
Let $A_1,\hdots,A_n$ denote the columns of $A,$ and let $(\xb,A) \in \C^n \times \C^{n^2}$ be such that 
$f^A(\xb) = \frac{\partial f^A(\xb)}{\partial X_{i+1}}=\hdots= \frac{\partial f^A(\xb)}{\partial X_n}=\bz.$
We want to show that the Jacobian of the system of polynomials 
\[
f(A\xb),(\grad(f)\cdot A_{i+1})(A\xb),\hdots, (\grad(f)\cdot A_n)(A\xb)
\]
at $(\xb,A)$ has full rank, which holds if and only if the Jacobian of the system of polynomials 
\[
f,\grad(f)\cdot A_{i+1},\hdots, \grad(f)\cdot A_n
\]
at $(\xb,A)$ has full rank. Let $\Psi_i$ be the mapping
%
\begin{align*}
    \Psi_i: \C^n \times \C^{n^2} &\rightarrow \C^{n-i+1} \\
       (\xb,A) &\mapsto 
       \left(f(\xb),\grad(f(\xb))\cdot 
A_{i+1},\hdots
\grad(f(\xb))\cdot 
A_{n}
\right).
\end{align*}
%
For all $(\xb,A) \in \Psi_i^{-1}(0),$ the Jacobian matrix of $\Psi_i$ at $(\xb,A)$ has full rank $n-i+1$. Indeed, set 
\[
F_j = \frac{\pa f}{\pa X_1} a_{i+j,1} + \hdots + \frac{\pa f}{\pa X_n}a_{i+j,n}, ~1 \leq j \leq n-i,
\]
so that the Jacobian is equal to
\[
\left[ 
\begin{array}{cccccc}
\frac{\pa f}{\pa X_1} \hdots \frac{\pa f}{\pa X_n}            &0\hdots 0&\hdots& 0 \hdots 0                                         & \hdots & 0\hdots 0 \\ 
\frac{\pa F_{1}}{\pa X_1}\hdots \frac{\pa F_{1}}{\pa X_n}     &0\hdots 0&\hdots& \frac{\pa f}{\pa X_1} \hdots \frac{\pa f}{\pa X_n} & \hdots & 0\hdots 0 \\
\ddots                                                        &0\hdots 0&\ddots&\ddots                                              & \ddots & \ddots \\
\frac{\pa F_{n-i}}{\pa X_1}\hdots \frac{\pa F_{n-i}}{\pa X_n} &0\hdots 0&\hdots&0 \hdots 0&  \hdots & \frac{\pa f}{\pa X_1} \hdots \frac{\pa f}{\pa X_n}  
\end{array}
\right]. 
\] 
Since $f$ is square-free, $I(V(f)) = \langle f \rangle.$ Therefore, for all $\xb \in V(f)$, 
\[
T_{\xb} V(f) = \textrm{Nullspace}(\grad_{\xb} f).
\]
Since $V(f)$ is smooth, $T_{\xb} V(f)$ has dimension $n-1$ and one partial vanishes. At least one partial of $f$ must be non-zero. Since  $\frac{\pa f}{\pa X_1} \hdots \frac{\pa f}{\pa X_n}$ appears in $n-i+1$ rows, the matrix has full rank $n-i+1$.
\end{proof}
%
Since $\deg \Phi_i \leq d,$ it follows by Proposition 3.1 that a hypersurface $\Xi_i \subset \C^{n^2}$ exists, of degree at most $d^{n+(n-i+1)}=d^{2n-i+1}\leq d^{2n},$ with the property that, if $A \in \C^{n^2}-\Xi_i,$ then 
%
\[
 \jac_{\xb} (\Phi_{i,A})=\jac_\xb \left(f^A,\frac{\pa f^A}{\pa X_{i+1}},\hdots,\frac{\pa f^A}{\pa X_n}\right)
\]
has full rank at all $\xb \in V(f^A)$, and therefore $f^A$ satisfies $\textbf{H}_i(1)$. Now, to prove the claim for all $i \in \{1,\hdots,n\},$ we take
\[
\Delta_1 := \bigcup_{i=1}^n \Xi_i,
\]
where we have the following bound on the degree which proves the claim:
\[
\deg \Delta_1 \leq \sum_{i=1}^n \deg \Xi_i  \leq d^{2n^2}.
\]

@@

\subsection{Application: property $\textbf{H}'_i$}
%
Let $f \in \ZZ[X_1,\hdots,X_n]$ have total degree $d$ with $V(f) \subset \C^n$ smooth. 
%
\begin{theorem}
  There exists a Zariski closed set $\Delta_2 \subset \C^{n-1}$ of
  degree at most $d^{2n^2}$ with the property that, if $\bm\alpha \in \C^{n-1} - \Delta_2$, then, for each $i \in \{1,\hdots,n\},$ if $f$ satisfies $\textbf{H}_i$ then $f$ and $\bm \alpha$ satisfy
  $\textbf{H}_i'$.
\end{theorem}
%
Fix $i \in \{1,\hdots,n\}$ and let $\Phi_i$ be the mapping 
%
\begin{align*}
    \Phi_i: \C^n \times \C^{i-1} &\rightarrow \C^{n} \\
           (\xb,\alpha_1,\hdots,\alpha_{i-1}) &\mapsto 
       (X_1(\xb)-\alpha_1,\hdots \\
       \hdots &,X_{i-1}(\xb)-\alpha_{i-1}, f(\xb),\frac{\pa f(\xb)}{\pa X_{i+1}},\hdots,\frac{\pa f(\xb)}{\pa X_n} ).
\end{align*}
%
\begin{proposition}
$\bz$ is a regular value of $\Phi_i.$
\end{proposition} 
\begin{proof}
At all $(\xb,\alpha_1,\hdots,\alpha_{i-1}) \in \Phi_i^{-1}(0),$ the Jacobian matrix of $\Phi_i$ has full rank. Indeed, the Jacobian is equal to
\[
\left[ 
\begin{array}{cc}
\hspace{-9mm}\bI_n\hspace{10mm}\hdots \hdots&\hspace{-10mm}-\bI_{i-1}\\
\jac_{\xb}\left(f,\frac{\pa
  f}{\pa X_{i+1}},\hdots,\frac{\pa f}{\pa X_n}\right)&\bz_{n-i+1\times i-1}
\end{array}
\right], 
\] 
and the Jacobian of $f,\frac{\pa
  f}{\pa X_{i+1}},\hdots,\frac{\pa f}{\pa X_n}$ at $\xb$ is never singular because we are assuming that $f$ satisfies $\textbf{H}_i$. Therefore the entire matrix must have full rank, and $\bz$ is a regular value of $\Phi_i.$
\end{proof}
%
Since $\deg \Phi_i \leq d,$ it follows by Proposition 3.1 that a hypersurface $\Xi_i \subset \C^{i-1}$ exists, of degree at most $d^{2n},$ with the property that, if $\bm \alpha_{i-1} \in \C^{i-1}-\Xi_i,$ then the Jacobian of $\Phi_{i,\alpha_{i-1}}$:
%
\[
 \jac_{\xb} \left(X_1(\xb)-\alpha_1,\hdots 
       ,X_{i-1}(\xb)-\alpha_{i-1}, f(\xb),\frac{\pa f(\xb)}{\pa X_{i+1}},\hdots,\frac{\pa f(\xb)}{\pa X_n}\right)
\]
has full rank at all $\xb \in V(f)$, and therefore $f$ and $\bm \alpha_{i-1}$ satisfy $\textbf{H}_i^{'}$. Now, to prove the claim for all $i \in \{1,\hdots,n\},$ we take
\[
\Delta_2 := \bigcup_{i=1}^n \Xi_i,
\]
where we have the following bound on the degree which proves the claim:
\[
\deg \Delta_2 \leq \sum_{i=1}^n \deg \Xi_i  \leq d^{2n^2}.
\]

@@

is integral.
%
%
%
%
\section{Proof of Theorem 3.1}
%
Recall Theorem 3.1: \textit{Suppose that $f \in \Q[X_1,\hdots,X_n]$ satisfies \textbf{H}. Then there exists a Zariski closed set $\D \subset \mathbb{C}^{n^2}$ of degree at most $D^{\textrm{\td}}$ such that, if $\textbf{A}$ is chosen from  $\C^{n^2}-\Delta$ ( $\Q^{n^2}-\Delta$?) then $f^A$ satisfies $\textbf{H}_i^{'}$ for all $i \in \{1,\hdots,n\}.$} 
%
\begin{proof}
Assume that $f\in \Q[X_1,\hdots,X_n]$ has degree at most $D$ and satisfies $\textbf{H}.$ By Theorem 5.1, there exists a Zariski closed set $\D_1\subset \C^{n^2}$ with the property that, if $A \in \C^{n^2}-\D_1$ then, for $i \in \{1,\hdots,n\},$ $f^A$ satisfies $\textbf{H}_i^{'}(1), \textbf{H}_i^{'}(2),$ and $\textbf{H}_i^{'}(3).$ Likewise, by Theorem 6.1, we obtain $\D_2 \subset \C^{n^2}$ with the property that if $A \in \C^{n^2}-\D_2$ then $f^A$ satisfies $\textbf{H}_i^{'}(4), \textbf{H}_i^{'}(5),$ and $\textbf{H}_i^{'}(6).$ By Theorem 7.1, $\D_3 \subset \C^{n^2}$ exists with $f^A$ satisfying $\textbf{H}_i^{'}(7)$ when $A \in \C^{n^2}-\D_3$. Clearly then, if 
\[
A \in \C^{n^2} - \bigcup_{i=1}^3 \D_i
\]
then $f^A$ satisfies  $\textbf{H}_i^{'}$ for all $i \in \{1,\hdots,n\}.$ Take 
\[
\Delta = \bigcup_{i=1}^3 \D_i,
\]
and note that 
\[
\deg \D \leq  \sum_{i=1}^3 \deg \D_i.
\]
\end{proof}
%
Recall Corollary 3.2: \textit{Fix $S \subset \mathbb{Q}$ with $|S| \geq \epsilon^{-1} D^{\textrm{\td}}$ and $\epsilon > 0$. Then for $A\in S^{n^2}$ chosen randomly, the probability that $f^A$ satisfies $H_i^{'}$ for all $i \in \{1,\hdots,n\}$ is at least $1-\epsilon.$}
%
\begin{proof}
\[
Pr[A \in \Delta] \leq \frac{\deg \Delta}{|S|} \leq \frac{D^{\td}}{\epsilon^{-1}D^{\td}} = \epsilon.
\]
\end{proof}
%
%
%
%
\section{Bit complexity}
%
Recall Theorem 1.1: Suppose that $f\in\ZZ[X_1\hdots,X_n]$ is squarefree, satisfies
  $\deg(f) \leq d, \htt(f) \leq s$, and that $V(f) \subset \C^n$ is
  smooth.  Assume also that $f$ is given by a straight-line program
  $\Gamma$ of size $L$ with integer constants of height at most $b.$

>>>>

\keywords{Real algebraic geometry, transversality, Noether
  position, bit complexity}

@@

Let $f \in \ZZ[X_1,\hdots,X_n]$ have total degree $d$ with $V(f) \subset \C^n$ smooth. 
%
\begin{theorem}
There exists a hypersurface $\Delta_1 \subset \C^{n^2}$ of degree at most $nd^{2n}$ with the property that, if $A \in \C^{n^2} - \Delta_1$ then, for each $i \in \{1,\hdots,n\},$ $f^A$ satisfies $\textbf{H}_i(1)$.
\end{theorem}

For each $i \in \{1,\hdots,n\}$, let $\Phi_i$ be the polynomial mapping 
\[
\Phi_i : (\xb,A) \mapsto \left(f^A(\xb), \frac{\partial f^A}{\partial X_{i+1}}(\xb),\hdots,\frac{\partial f^A}{\partial X_n}(\xb)\right),
\] 
so that $\Phi_{i,A}^{-1}(0)= W(\pi_i,V^A).$
%
\begin{proposition} 
$\bz$ is a regular value of $\Phi_i$.
\end{proposition}
%
\begin{proof}
Let $A_1,\hdots,A_n$ denote the columns of $A,$ and let $(\xb,A) \in \C^n \times \C^{n^2}$ be such that 
$f^A(\xb) = \frac{\partial f^A(\xb)}{\partial X_{i+1}}=\hdots= \frac{\partial f^A(\xb)}{\partial X_n}=\bz.$
We want to show that the Jacobian of the system of polynomials 
\[
f(A\xb),(\grad(f)\cdot A_{i+1})(A\xb),\hdots, (\grad(f)\cdot A_n)(A\xb)
\]
at $(\xb,A)$ has full rank, which holds if and only if the Jacobian of the system of polynomials 
\[
f,\grad(f)\cdot A_{i+1},\hdots, \grad(f)\cdot A_n
\]
at $(\xb,A)$ has full rank. Let $\Psi_i$ be the mapping
%
\begin{align*}
    \Psi_i: \C^n \times \C^{n^2} &\rightarrow \C^{n-i+1} \\
       (\xb,A) &\mapsto 
       \left(f(\xb),\grad(f(\xb))\cdot 
A_{i+1},\hdots
\grad(f(\xb))\cdot 
A_{n}
\right).
\end{align*}
%
For all $(\xb,A) \in \Psi_i^{-1}(0),$ the Jacobian matrix of $\Psi_i$ at $(\xb,A)$ has full rank $n-i+1$. Indeed, set 
\[
F_j = \frac{\pa f}{\pa X_1} a_{i+j,1} + \hdots + \frac{\pa f}{\pa X_n}a_{i+j,n}, ~1 \leq j \leq n-i,
\]
so that the Jacobian is equal to
\[
\left[ 
\begin{array}{cccccc}
\frac{\pa f}{\pa X_1} \hdots \frac{\pa f}{\pa X_n}            &0\hdots 0&\hdots& 0 \hdots 0                                         & \hdots & 0\hdots 0 \\ 
\frac{\pa F_{1}}{\pa X_1}\hdots \frac{\pa F_{1}}{\pa X_n}     &0\hdots 0&\hdots& \frac{\pa f}{\pa X_1} \hdots \frac{\pa f}{\pa X_n} & \hdots & 0\hdots 0 \\
\ddots                                                        &0\hdots 0&\ddots&\ddots                                              & \ddots & \ddots \\
\frac{\pa F_{n-i}}{\pa X_1}\hdots \frac{\pa F_{n-i}}{\pa X_n} &0\hdots 0&\hdots&0 \hdots 0&  \hdots & \frac{\pa f}{\pa X_1} \hdots \frac{\pa f}{\pa X_n}  
\end{array}
\right]. 
\] 
Since $f$ is square-free, $I(V(f)) = \langle f \rangle.$ Therefore, for all $\xb \in V(f)$, 
\[
T_{\xb} V(f) = \textrm{Nullspace}(\grad_{\xb} f).
\]
Since $V(f)$ is smooth, $T_{\xb} V(f)$ has dimension $n-1$ and one partial vanishes. At least one partial of $f$ must be non-zero. Since  $\frac{\pa f}{\pa X_1} \hdots \frac{\pa f}{\pa X_n}$ appears in $n-i+1$ rows, the matrix has full rank $n-i+1$.
\end{proof}
%
Since $\deg \Phi_i \leq d,$ it follows by Proposition 3.1 that a hypersurface $\Xi_i \subset \C^{n^2}$ exists, of degree at most $d^{n+(n-i+1)}=d^{2n-i+1}\leq d^{2n},$ with the property that, if $A \in \C^{n^2}-\Xi_i,$ then 
%
\[
 \jac_{\xb} (\Phi_{i,A})=\jac_\xb \left(f^A,\frac{\pa f^A}{\pa X_{i+1}},\hdots,\frac{\pa f^A}{\pa X_n}\right)
\]
has full rank at all $\xb \in V(f^A)$, and therefore $f^A$ satisfies $\textbf{H}_i(1)$. Now, to prove the claim for all $i \in \{1,\hdots,n\},$ we take
\[
\Delta_1 := \bigcup_{i=1}^n \Xi_i,
\]
where we have the following bound on the degree which proves the claim:
\[
\deg \Delta_1 \leq \sum_{i=1}^n \deg \Xi_i  \leq nd^{2n}.
\]

@@

\subsection{Application: property $\textbf{H}'_i$}
%
Let $f \in \ZZ[X_1,\hdots,X_n]$ have total degree $d$ with $V(f) \subset \C^n$ smooth. 
%
\begin{theorem}
  There exists a Zariski closed set $\Delta_2 \subset \C^{n-1}$ of
  degree at most $nd^{2n}$ with the property that, if $\bm\alpha \in \C^{n-1} - \Delta_2$, then, for each $i \in \{1,\hdots,n\},$ if $f$ satisfies $\textbf{H}_i$ then $f$ and $\bm \alpha$ satisfy
  $\textbf{H}_i'$.
\end{theorem}
%
Fix $i \in \{1,\hdots,n\}$ and let $\Phi_i$ be the mapping 
%
\begin{align*}
    \Phi_i: \C^n \times \C^{i-1} &\rightarrow \C^{n} \\
           (\xb,\alpha_1,\hdots,\alpha_{i-1}) &\mapsto 
       (X_1(\xb)-\alpha_1,\hdots \\
       \hdots &,X_{i-1}(\xb)-\alpha_{i-1}, f(\xb),\frac{\pa f(\xb)}{\pa X_{i+1}},\hdots,\frac{\pa f(\xb)}{\pa X_n} ).
\end{align*}
%
\begin{proposition}
$\bz$ is a regular value of $\Phi_i.$
\end{proposition} 
\begin{proof}
At all $(\xb,\alpha_1,\hdots,\alpha_{i-1}) \in \Phi_i^{-1}(0),$ the Jacobian matrix of $\Phi_i$ has full rank. Indeed, the Jacobian is equal to
\[
\left[ 
\begin{array}{cc}
\hspace{-9mm}\bI_n\hspace{10mm}\hdots \hdots&\hspace{-10mm}-\bI_{i-1}\\
\jac_{\xb}\left(f,\frac{\pa
  f}{\pa X_{i+1}},\hdots,\frac{\pa f}{\pa X_n}\right)&\bz_{n-i+1\times i-1}
\end{array}
\right], 
\] 
and the Jacobian of $f,\frac{\pa
  f}{\pa X_{i+1}},\hdots,\frac{\pa f}{\pa X_n}$ at $\xb$ is never singular because we are assuming that $f$ satisfies $\textbf{H}_i$. Therefore the entire matrix must have full rank, and $\bz$ is a regular value of $\Phi_i.$
\end{proof}
%
Since $\deg \Phi_i \leq d,$ it follows by Proposition 3.1 that a hypersurface $\Xi_i \subset \C^{i-1}$ exists, of degree at most $d^{2n},$ with the property that, if $\bm \alpha_{i-1} \in \C^{i-1}-\Xi_i,$ then the Jacobian of $\Phi_{i,\alpha_{i-1}}$:
%
\[
 \jac_{\xb} \left(X_1(\xb)-\alpha_1,\hdots 
       ,X_{i-1}(\xb)-\alpha_{i-1}, f(\xb),\frac{\pa f(\xb)}{\pa X_{i+1}},\hdots,\frac{\pa f(\xb)}{\pa X_n}\right)
\]
has full rank at all $\xb \in V(f)$, and therefore $f$ and $\bm \alpha_{i-1}$ satisfy $\textbf{H}_i^{'}$. Now, to prove the claim for all $i \in \{1,\hdots,n\},$ we take
\[
\Delta_2 := \bigcup_{i=1}^n \Xi_i,
\]
where we have the following bound on the degree which proves the claim:
\[
\deg \Delta_2 \leq \sum_{i=1}^n \deg \Xi_i \leq n d^{2n}.
\]

@@

is integral.
%
%
%
%
\section{Proof of Theorem 2.1}
%
Recall Theorem 3.1: \textit{Suppose that $f \in \Q[X_1,\hdots,X_n]$ satisfies \textbf{H}. Then there exists a Zariski closed set $\D \subset \mathbb{C}^{n^2}$ of degree at most $D^{\textrm{\td}}$ such that, if $\textbf{A}$ is chosen from  $\C^{n^2}-\Delta$ ( $\Q^{n^2}-\Delta$?) then $f^A$ satisfies $\textbf{H}_i^{'}$ for all $i \in \{1,\hdots,n\}.$} 
%
\begin{proof}
Assume that $f\in \Q[X_1,\hdots,X_n]$ has degree at most $D$ and satisfies $\textbf{H}.$ By Theorem 5.1, there exists a Zariski closed set $\D_1\subset \C^{n^2}$ with the property that, if $A \in \C^{n^2}-\D_1$ then, for $i \in \{1,\hdots,n\},$ $f^A$ satisfies $\textbf{H}_i^{'}(1), \textbf{H}_i^{'}(2),$ and $\textbf{H}_i^{'}(3).$ Likewise, by Theorem 6.1, we obtain $\D_2 \subset \C^{n^2}$ with the property that if $A \in \C^{n^2}-\D_2$ then $f^A$ satisfies $\textbf{H}_i^{'}(4), \textbf{H}_i^{'}(5),$ and $\textbf{H}_i^{'}(6).$ By Theorem 7.1, $\D_3 \subset \C^{n^2}$ exists with $f^A$ satisfying $\textbf{H}_i^{'}(7)$ when $A \in \C^{n^2}-\D_3$. Clearly then, if 
\[
A \in \C^{n^2} - \bigcup_{i=1}^3 \D_i
\]
then $f^A$ satisfies  $\textbf{H}_i^{'}$ for all $i \in \{1,\hdots,n\}.$ Take 
\[
\Delta = \bigcup_{i=1}^3 \D_i,
\]
and note that 
\[
\deg \D \leq  \sum_{i=1}^3 \deg \D_i.
\]
\end{proof}
%
Recall Corollary 3.2: \textit{Fix $S \subset \mathbb{Q}$ with $|S| \geq \epsilon^{-1} D^{\textrm{\td}}$ and $\epsilon > 0$. Then for $A\in S^{n^2}$ chosen randomly, the probability that $f^A$ satisfies $H_i^{'}$ for all $i \in \{1,\hdots,n\}$ is at least $1-\epsilon.$}
%
\begin{proof}
\[
Pr[A \in \Delta] \leq \frac{\deg \Delta}{|S|} \leq \frac{D^{\td}}{\epsilon^{-1}D^{\td}} = \epsilon.
\]
\end{proof}
%
%
%
%
\section{Bit complexity}
%
Recall Theorem 1.1: Suppose that $f\in\ZZ[X_1\hdots,X_n]$ is squarefree, satisfies
  $\deg(f) \leq d, \htt(f) \leq s$, and that $V(f) \subset \C^n$ is
  smooth.  Assume also that $f$ is given by a straight-line program
  $\Gamma$ of size $L$ with integer constants of height at most $b.$
